---
title: "Results comparing smoothing first and smoothing during the clustering procedure"
output: html_document
---

We compare two different approaches: the original incorporates the smoothing as part of the clustering algorithm, while the alternative smoothes the data first and then performs clustering.

```{r}
path = "D:/Grad School/projects/sequence_clustering"
setwd(path)
load("results/analysis_oas1/res_initial_smooth_3.RData")
res_priorsmooth = res
load("results/analysis_oas1/res_smooth_3.RData")
res_smooth = res
```


We first plot what the clusters from the two approaches look like
```{r}
#plot the clusters
par(mfrow = c(2, 1))
plot(res_priorsmooth$phi[1, ], col = 2, ylim = c(0, 0.015), type = 'l', main = "unsmoothed clusters")
lines(res_priorsmooth$phi[2, ], col = 3)
lines(res_priorsmooth$phi[3, ], col = 4)
plot(res_smooth$phi[1, ], col = 3, ylim = c(0, 0.015), type = 'l', main = "smoothed clusters")
lines(res_smooth$phi[2, ], col = 2)
lines(res_smooth$phi[3, ], col = 4)
```


```{r}
load(paste0(path, "/data/oas1/OAS1.Robj"))
comp_priorsmooth = cbind(res_priorsmooth$pi, Robj$g)

#compare the actual genotype with the estimated profile
est_priorsmooth = res_priorsmooth$pi %*% res_priorsmooth$lambda
est_smooth = res_smooth$pi %*% res_smooth$lambda
```

We then choose three examples, one from each genotype class, and look at what the reconstructed signal looks like based on the clusters.

```{r}
#look at example profiles belonging to different genotype classes
par(mfrow = c(2, 1))
plot(est_priorsmooth[48, ], ylim = c(0, 10), type = 'l', main = "estimated signals, unsmoothed")
lines(est_priorsmooth[17, ], col = 2)
lines(est_priorsmooth[3, ], col = 3)
plot(est_smooth[48, ], ylim = c(0, 10), type = 'l', main = "estimated signals with smoothing procedures")
lines(est_smooth[17, ], col = 2)
lines(est_smooth[3, ], col = 3)
```

And finally, smooth the unsmoothed signals post hoc and add in signals estimated from the original data

```{r}
#smooth unsmoothed profiles post-hoc
library(smash)
par(mfrow = c(3, 1))
plot(ashsmooth.pois(data[48, 2049:4096], log = FALSE), ylim = c(0, 7), type = 'l', main = "original data, smoothed post hoc")
lines(ashsmooth.pois(data[17, 2049:4096], log = FALSE), col = 2)
lines(ashsmooth.pois(data[3, 2049:4096], log = FALSE), col = 3)
plot(est_priorsmooth[48, ], ylim = c(0, 7), type = 'l', main = "estimated signals, smoothed prior to clustering")
lines(est_priorsmooth[17, ], col = 2)
lines(est_priorsmooth[3, ], col = 3)
plot(est_smooth[48, ], ylim = c(0, 7), type = 'l', main = "estimated signals with smoothing procedure")
lines(est_smooth[17, ], col = 2)
lines(est_smooth[3, ], col = 3)
```