---
title: "structure plots and heat maps for K = 4"
output: html_document
---

---
title: "OAS1 - structure plots, heat maps and further exploration"
output: html_document
---

We first look at the structure plot


```{r}
library(gplots)
library(smash)
path = "D:/Grad School/projects/sequence_clustering"
setwd(path)
load("results/analysis_oas1/res_nosmooth_4.RData")
res_nosmooth = res
load("results/analysis_oas1/res_smooth_4.RData")
res_smooth = res
load(paste0(path, "/data/oas1/OAS1.Robj"))
```


```{r}
K = dim(res_nosmooth$pi)[2]
nosmooth_pi_unordered = res_nosmooth$pi
ordering = order(Robj$g)
nosmooth_pi_ordered = nosmooth_pi_unordered[ordering, ]
smooth_pi_unordered = res_smooth$pi
smooth_pi_ordered = smooth_pi_unordered[ordering, ]
sep_lines = cumsum(table(Robj$g))[1:(K - 1)]

par(mfrow = c(2, 1), mar=c(2, 2, 2, 2), oma = c(4, 4, 0.2, 0.2))
barplot(t(nosmooth_pi_ordered), col = 2:(K + 1), axisnames = F, space = 0, border = NA, main = "structure plot for unsmoothed model", las = 1, ylim = c(0, 1), cex.axis = 1.5, cex.main = 1.4)
abline(v = sep_lines, lwd = 2)
barplot(t(smooth_pi_ordered), col = c(3, 2, 4:(K + 1)), axisnames = F, space = 0, border = NA, main = "structure plot for smoothed model", las = 1, ylim = c(0, 1), cex.axis = 1.5, cex.main = 1.4)
abline(v = sep_lines, lwd = 2)
```


Now we take a look at the heatmap for genotypes 1 and 2. First heatmap shows correlation for raw reads, second shows correlation for admixture proportions using the model with no smoothing, and the third shows correlation for admixture proportions using the model with smoothing.

```{r}
data_subset = data[Robj$g!=0, 2049:4096]
g_subset = Robj$g[Robj$g != 0]

col = c(rgb(seq(0, 1, length = 15), 1, seq(0, 1, length = 15)), rgb(1, seq(1, 0 ,length = 15), seq(1, 0, length = 15)))

heatmap.2(cor(t(data_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))


nosmooth_subset = res_nosmooth$pi[Robj$g!=0, ]

hm_nosmooth_12 = heatmap.2(cor(t(nosmooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

smooth_subset = res_smooth$pi[Robj$g!=0, ]

heatmap.2(cor(t(smooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

```

The conclusions are generally similar to the results from having 3 clusters. The model with smoothing now shows more dissimilarities between samples in genotype classes 1 and 2 and less clustering, which is to be expected.

Now we draw the same heatmaps, but for genotypes 0 and 2.

```{r}
data_subset = data[Robj$g!=1, 2049:4096]
g_subset = Robj$g[Robj$g != 1]

col = c(rgb(seq(0, 1, length = 15), 1, seq(0, 1, length = 15)), rgb(1, seq(1, 0 ,length = 15), seq(1, 0, length = 15)))

heatmap.2(cor(t(data_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))


nosmooth_subset = res_nosmooth$pi[Robj$g!=1, ]

hm_nosmooth_02 = heatmap.2(cor(t(nosmooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

smooth_subset = res_smooth$pi[Robj$g!=1, ]

hm_smooth_02 = heatmap.2(cor(t(smooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

```

Again, conclusions are similar. Model with smoothing still distinguishes between genotype classes 0 and 2 clearly, though model without smoothing has a sample from genotype 2 that is clustered together with 0.


Finally, for genotypes 0 and 1

```{r}
data_subset = data[Robj$g!=2, 2049:4096]
g_subset = Robj$g[Robj$g != 2]

col = c(rgb(seq(0, 1, length = 15), 1, seq(0, 1, length = 15)), rgb(1, seq(1, 0 ,length = 15), seq(1, 0, length = 15)))

heatmap.2(cor(t(data_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))


nosmooth_subset = res_nosmooth$pi[Robj$g!=2, ]

hm_nosmooth_01 = heatmap.2(cor(t(nosmooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

smooth_subset = res_smooth$pi[Robj$g!=2, ]

hm_smooth_01 = heatmap.2(cor(t(smooth_subset), method = "pearson"), labCol = g_subset, labRow = g_subset, scale = "none", trace = "none", distfun = function(x) dist(x, method = "euclidean"), col = col, hclustfun = function(x) hclust(x,method="average"), breaks = c(seq(-1, 0, 0.1), seq(0.05, 1, length.out = 20)))

```

Again, similar conclusions, except for that one sample from genotype 0 that is clustered into 1 using the model with smoothing and more dissimilarities in general.

