---
title: "Plot results from clustering analysis"
output: html_document
---


```{r}
gene_region = read.table("../../supplemental/gtex/gene_region.txt", stringsAsFactors = FALSE)
gene_region = unlist(gene_region, use.names = FALSE)
gene_region = gsub("\\:", "_", gene_region)
gene_region = gsub("\\-", "_", gene_region)
```


```{r, echo=FALSE, fig.height = 10, fig.width = 8}


# for(K in 2:5){
#   for(i in 1:length(gene_region)){
#     load_name = paste("cluster_res_100", K, "clusters", gene_region[i], sep = "_")
#     load_name = paste0(load_name, ".RData")
#     if(!file.exists(load_name)) next
#     load(load_name)
#     
#     n = dim(res$pi)[1]
#     sep_lines = seq(0, n, length.out = 9)
#     sep_lines = sep_lines[c(-1, -length(sep_lines))]
#     
#     par(mfrow = c(2, 1))
#     plot(res$phi[1, ], type = 'l', col = 2, main = paste0("cluster center for ", region))
#     for(k in 2:K){
#       lines(res$phi[k, ], col = k + 1)
#     }
#     barplot(t(res$pi), col = 2:(K + 1), axisnames = F, space = 0, border = NA, main = paste0("structure plot for ", region), las = 1, ylim = c(0, 1), cex.axis = 1.5, cex.main = 1.4)
#     abline(v = sep_lines, lwd = 2)
#   }
# }


```{r, echo=FALSE, fig.height = 10, fig.width = 8}


for(K in 2:5){
  for(i in 1:length(gene_region)){
    load_name = paste("cluster_res_all", K, "clusters", gene_region[i], sep = "_")
    load_name = paste0(load_name, ".RData")
    if(!file.exists(load_name)) next
    load(load_name)
    
    sep_lines = 0
    for(j in 1:(length(reads) - 1)){
      temp = dim(reads[[j]][[2]])[1]
      sep_lines[j + 1] = sep_lines[j] + temp
    }
    
    sep_lines = sep_lines[c(-1, -length(reads))]

    tissue_name_full = NULL
    for(j in 1:(length(reads) - 1)){
      temp = rep(tissue_name[j], dim(reads[[j]][[2]])[1])
      tissue_name_full = c(tissue_name_full, temp)
    }
    
    par(mfrow = c(2, 1))
    plot(res$phi[1, ], type = 'l', col = 2, main = paste0("cluster center for ", region))
    for(k in 2:K){
      lines(res$phi[k, ], col = k + 1)
    }
    barplot(t(res$pi), col = 2:(K + 1), axisnames = F, space = 0, border = NA, main = paste0("structure plot for ", region), las = 1, ylim = c(0, 1), cex.axis = 1.5, cex.main = 1.4)
    abline(v = sep_lines, lwd = 2)
  }
}
```


```

